---
title: "Portfolio"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


---

*Welcome!*

I'm Aaron Begg, a professional in business communications both internally and client-facing. As a lifelong learner I've recently expanded my skill set through the use of data visualization and analytics. 
Developing reports with R and RStudio has given me exciting new tools to work with data, These tools help me identify statistical relationships and create effective visuals and reports (such as this site) that clearly represent the data.



## What I Do

I specialize in:

> - 📊 Data Analysis: Cleaning, transforming, and analyzing business data
> - 📈 Visualization: Creating clear, compelling charts with ggplot2
> - 📝 Reporting: Building professional reports that tell data stories
> - 🔍 Problem Solving: Applying statistical thinking to real-world
challenges


### Sample Work


Here's an example of my visualization style:



```{r sample}

sample_plot <- readRDS("docs/scatter_plot.rds")
sample_plot

```

## Portfolio Highlights

Browse my projects to see:

[Government Data Analysis](project2.html): Working with public datasets to extract
business insights

[Performance Analytics](project1.html): Analyzing operational data for decision-making
Statistical Exploration: Uncovering patterns in real-world data

## Technical Skills

Tools: R, RStudio, R Markdown, ggplot2, dplyr
Analysis: Data cleaning, exploratory analysis, statistical visualization
Communication: Report generation, data storytelling, executive summaries

## Let's Connect

- ✉️ Email: [aaronfraserbegg@gmail.com](mailto:aaronfraserbegg@gmail.com)
- 🌐 LinkedIn: [Aaron Begg](https://www.linkedin.com/in/aaron-begg-7b92a82a1/)
- 👾 GitHub: [aaronbegg](https://github.com/aaronbegg)

This portfolio showcases projects from my MGMT coursework and
demonstrates practical skills in R and statistical analysis.