---
title: "Data Import and Exploration"
subtitle: "CSV Import and Data Cleaning"
author: "Aaron Begg -- 567 070 685"
date: "`r Sys.Date()`"
site: my-portfolio
---

# Data Import and Exploration

```{r setup}
regional_sales <- data.frame(
  region = c("North", "South", "East", "West", "North", "South", "East", "West",
             "North", "South", "East", "West", "North", "South", "East", 
             "North", "South","North"),  
  quarter = c(rep(c("Q1", "Q2", "Q3", "Q4"), each = 4), "Q4", "Q4"),
  revenue = c(125000, 98000, 115000, 102000, 
              138000, 105000, 128000, 118000,
              142000, 89000, 135000, 125000,
              155000, 112000, 148000, 165000, 
              120000, 91250),
  customers = c(450, 380, 420, 395, 
                485, 415, 465, 445,
                520, 350, 485, 470,
                575, 425, 545, 590, 435,324)
)

```

## a.

```{r eval=TRUE, echo=TRUE}
str(regional_sales)
```
> There are 18 observations of 4 variables. The data types of the variables are:

- Region (chr)
- Quarter (chr)
- Revenue (num)
- Customers (num)

> For 4 quarters of data between 4 regions, we would expect only 16 data points. Further investigation may be needed to find out if these were partial data entries that need to be combined with similar data, or erroneously completed entries that should be removed from the dataset.

## b.

```{r eval=TRUE, echo=TRUE}
head(regional_sales)
```

> The first value shown in the table is for the North Region.


```{r eval=TRUE, echo=TRUE}
summary(regional_sales)
```

> The range of revenue values is $89000 to $165000.

## c.

```{r eval=TRUE, echo=TRUE}
unique(regional_sales)
```

> The dataset provides 18 unique combinations of 4 Regions and 4 Quarters, meaning we have extra data within our set. 

## d.

```{r eval=TRUE, echo=TRUE}
sum(regional_sales$customers)
```

> This total customer count gives us the ability to apply metrics to each Quarter and Region so we can compare performance.

# Data Quality Assessment

## a.

```{r eval=TRUE, echo=TRUE}
sum(is.na(regional_sales))
```

> There are 0 missing values in the dataset. Missing values can skew the results of our data or may be important information that we have to manually find.

## b.

```{r eval=TRUE, echo=TRUE}
table(regional_sales$region, regional_sales$quarter)
```

> Records for each combination of Region and Quarter show that there is a missing data entry for Region West in Quarter 4. There are also 2 extra datasets for Region North in Q4, and 1 extra dataset for Region South, in Q4. This suggests there may be a combination of misreported data as well as extra datapoints that should either be combined or scrubbed from the dataset.

## c.

```{r eval=TRUE, echo=TRUE}
regional_sales[regional_sales$revenue < 90000, ]
```

> During Quarter 3, the South Region showed concerning performance trends with less than $90000 revenue.

# Business Logic Analysis

## a.

```{r eval=TRUE, echo=TRUE}
high_performance <- regional_sales[regional_sales$revenue > 140000, ]

high_performance[, c("region", "quarter", "revenue", "customers")]

nrow(high_performance)

```

> The Northern Region showed the majority of the highest revenues, and similarly the data shows Quarter 4 generating the majority of high revenue results. There were 4 high-performance records (revenues > $140000). This data is skewed due to the additional Q4 entries in the dataset.

## b.

```{r eval=TRUE, echo=TRUE}
regional_sales[regional_sales$quarter == "Q4" & regional_sales$revenue > 130000 & regional_sales$customers > 500, ]
```

> Two of the three months that had high revenue and traffic ( traffic > 500 and revenue > $130000) were in the North Region. This data is useful in showing regions with high conversion rates on customers, generating large profits from traffic, but is also skewed by the 2 additional data entries for North Region in Q4.

## c.

```{r eval=TRUE, echo=TRUE}
north_east <- regional_sales[regional_sales$region == "North" | regional_sales$region == "East", ]

north_east[, c("region", "quarter", "revenue", "customers")]

mean(north_east$revenue)
```

## d.

```{r eval=TRUE, echo=TRUE}
performance_flag <- regional_sales[regional_sales$revenue < 100000 | regional_sales$customers < 400, ]

performance_flag[, c("region", "quarter", "revenue", "customers")]

pf_count <- nrow(performance_flag)

rs_count <- nrow(regional_sales)

percentage <- (pf_count / rs_count) * 100

percentage
```

> Quarters where revenue > $100000 and customers > 400 accounted for 4 of the 18 records, or 22% of business quarters.

# Data Aggregation and Business Insights

## a.

```{r eval=TRUE, echo=TRUE}
aggregate(revenue ~ region, data = regional_sales, sum)
```

> The highest revenues were generated by the North Region, valued at $816250. 

## b.

```{r eval=TRUE, echo=TRUE}
aggregate(customers ~ quarter, data = regional_sales, mean)
```

> Quarter 4 shows the highest average custommer count. This may be explained by seasonal buying habits or a more successful marketing campaign running in Quarter 4, but the most likely answer is the 2 additional datapoints which are skewing the results.

# Executive Summary Creation

## a.

```{r eval=TRUE, echo=TRUE}
rev_total <- sum(regional_sales$revenue)

rev_q1 <- regional_sales[regional_sales$quarter == "Q1", ]
rev_q2 <- regional_sales[regional_sales$quarter == "Q2", ]
rev_q3 <- regional_sales[regional_sales$quarter == "Q3", ]
rev_q4 <- regional_sales[regional_sales$quarter == "Q4", ]

high_performance <- regional_sales[regional_sales$revenue > 140000, ]

rev_total
means <- c(
mean(rev_q1$revenue),
mean(rev_q2$revenue),
mean(rev_q3$revenue),
mean(rev_q4$revenue)
)

quarters <- c("Q1", "Q2", "Q3", "Q4")

bars <- barplot(
  means, 
  names.arg = quarters, 
  main = "Average Revenue by Quarter",
  xlab = "Quarter",
  ylab = "Average Revenue ($)",
  col = c("skyblue", "lightgreen", "orange", "plum")
)
text(bars, means, labels = round(means, 0), pos = 1)


high_performance


```

## b.

> Based off of the sales data provided, we see consistent performances in the North Region. 3 of the top 4 performing quarters were from the North Region, with $462000 total revenue in its top-3 quarters.

> Quarterly-based, performance is consistently higher in Q4 than other quarters. Quarter 4 had the highest average revenue at $131875, and also had 3 of the 4 highest revenue quarters overall.

---

[← Back to Projects](projects.html) | [← Back to Home](index.html)

